---
title: "spatial-data-intro"
author: "Clara Feigelson"
date: "`r Sys.Date()`"
output: html_document
---

# Intro to Spatial Data in R (Day 2)

```{r}
source("setup.R")

library(colorspace)
```

### 2.1 Vector Data

```{r Polygons}
# Copied from lesson plan

# download county shapefile for the state of Colorado
counties <- counties(state = "CO")
```

```{r Lines}
# Copied from lesson plan

roads <- roads(state = "CO", county = "Larimer")
```

```{r tmap}
# Copied from lesson plan

tmap_mode("view")

#Using qtm
qtm(counties)+
  qtm(roads)

#Using tm_shape
tm_shape(counties)+
  tm_polygons()+
tm_shape(roads)+
  tm_lines()
```

```{r sf}
# Copied from lesson plan

# filter out only Poudre Canyon Highway road data
poudre_hwy <- roads %>% 
  filter(FULLNAME == "Poudre Canyon Hwy")

qtm(poudre_hwy)
```

```{r Points}
# copied from lesson plan

poudre_points <- data.frame(name = c("Mishawaka", "Rustic", "Blue Lake Trailhead"),
                            long = c(-105.35634, -105.58159, -105.85563),
                            lat = c(40.68752, 40.69687, 40.57960))

poudre_points_sf <- st_as_sf(poudre_points, coords = c("long", "lat"), crs = 4326)

qtm(poudre_hwy)+
  qtm(poudre_points_sf)

```

```{r CRS}
# Copied from lesson plan

# see the CRS in the header metadata:
counties

#return just the CRS (more detailed)
st_crs(counties)

# check if two objects have same CRS
st_crs(counties) == st_crs(poudre_points_sf)

# transform one object's CRS to another's CRS
poudre_points_prj <- st_transform(poudre_points_sf, st_crs(counties))

#Now check that they match
st_crs(poudre_points_prj) == st_crs(counties)

```

### 2.2 Raster Data

```{r elevatr}
# Copied from lesson plan

elevation <- get_elev_raster(counties, z = 7)

qtm(elevation)

tm_shape(elevation)+
  tm_raster(style = "cont", title = "Elevation (m)")

elevation
```

```{r terra}
# Copied from lesson plan

# If we try this, we get an error
# elevation_crop <- crop(elevation, ext(poudre_hwy))

# converting rasterlayer object to a terra SpatRaster object
elevation <- rast(elevation)
names(elevation) <- "Elevation"

# run terra functions
elevation_crop <- crop(elevation, ext(roads))

tm_shape(elevation_crop)+
  tm_raster(style = "cont")

# final map
tm_shape(elevation, bbox = st_bbox(poudre_hwy))+
  tm_raster(style = "cont", title = "Elevation (m)")+
tm_shape(poudre_hwy)+
  tm_lines()+
tm_shape(poudre_points_prj)+
  tm_dots(size = 0.2)
```

### 3.1 Writing spatial data

```{r save vector data}
write_sf(poudre_hwy, "data/poudre_hwy.shp")

write_sf(poudre_points_prj, "data/poudre_points.shp")

```

```{r save raster data}
writeRaster(elevation_crop, "data/elevation_larimer.tif")
```

### 3.2 .RData files

```{r save .RData file}
save(poudre_hwy, poudre_points_prj, file = "data/spatial_objects.RData")
```

```{r remove and load files}
rm(poudre_hwy, poudre_points_prj)

load("data/spatial_objects.RData")
saveRDS(elevation_crop, "data/elevation_crop.RDS")
readRDS("data/elevation_crop.RDS") %>% rast()
```

### 3.3 Reading Spatial Data

```{r read shape files}
read_sf("data/poudre_hwy.shp")
```

```{r read raster files}
rast("data/elevation_larimer.tif")
```

## 4. Exercises

```{r Q1}
sub_counties <- counties %>%
  filter(NAME %in% c("Larimer","Denver","Pueblo"))

```

```{r Q2}

tm_shape(counties)+
  tm_polygons(col = "ALAND")

tm_shape(counties)+
  tm_polygons(col = "AWATER")


```

```{r Q3}
poudre_points_prj$elevation<-extract(elevation, poudre_points_prj$Elevation)

poudre_points_prj %>%
ggplot()+
  geom_bar(mapping = aes(x = name, y = elevation), stat = "identity")
```

Q4.

Perhaps there are 4 features of the Poudre Canyon Highway because it is not a straight line and is made up of smaller sections of a single highway. The linestrings seem to have different coordinates, which could be the geometry of the smaller sections.
